---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggflower

<!-- badges: start -->
[![R-CMD-check](https://github.com/srvanderplas/ggflower/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/srvanderplas/ggflower/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/srvanderplas/ggflower/branch/main/graph/badge.svg?token=tFnnmH1WMi)](https://codecov.io/gh/srvanderplas/ggflower)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Documentation](https://github.com/srvanderplas/ggflower/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/srvanderplas/ggflower/actions/workflows/pkgdown.yaml)
[![Test Coverage Workflow](https://github.com/srvanderplas/ggflower/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/srvanderplas/ggflower/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

The goal of ggflower is to provide flower-like charts using the grammar of graphics API in ggplot2. These charts are intended primarily to draw viewers in, rather than to accurately display data, as we know that radar charts (which are the broader class of charts that encompass flower charts) are harder to use to identify differences and make comparisons. 

## Installation

You can install the development version of ggflower from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("drbradford12/ggflower")
```

## Example

`ggflower` works by creating a "petal" shape for each categorical x variable and then using polar coordinates to turn these values into a flower shape. 

```{r example}
library(ggplot2)
library(ggflower)
## basic example code
n <- 4
dframe <- data.frame(index = 1:n, y = n:1)

# don't do this - use polar coordinates
dframe |>
  ggplot(aes(x = index, y = y)) +
  geom_flower(aes(fill = factor(index)), shape="heart") 


dframe |>
  ggplot(aes(x = index, y = y)) +
  geom_flower(aes(fill = factor(index)), shape="heart") +
  coord_polar()
```

As a result, it is critical that you use `coord_polar()` with `ggflower` plots. 


`ggflower` works well with facets and other `ggplot2` geoms.
```{r example2}

n <- 12
dframe <- data.frame(index = 1:n, y = abs(rnorm(n, mean=2)))

dframe |>
  ggplot(aes(x = index, y = y)) +
  geom_flower(aes(fill = y), shape="circle", colour = NA) +
  coord_polar() +
  theme_void() +
  scale_fill_gradient(low="yellow", high="darkorange")


n <- 12
k <- 9
dframe <- data.frame(
  expand.grid(
  index = 1:n, facet=1:k), y = abs(rnorm(n*k, mean=2)))

dframe |>
  ggplot(aes(x = index, y = y)) +
  geom_flower(shape="circle", aes(fill = y), colour=NA) +
  facet_wrap(~facet) +
  coord_polar() +
  theme_void() +
  scale_fill_gradient(low="yellow", high="darkorange") +
  theme(legend.position = "none")
```

There are several different petal types available:
```{r example3, out.width = "19%", fig.width = 2, fig.height = 2}
set.seed(3425)
n <- 12
dframe <- data.frame(index = 1:n, y = abs(rnorm(n, mean = 2)))

shapes <- c("box", "cos", "normal", "circle", "heart")
for (i in shapes) {
  dframe |>
    (\(df) ggplot(df, aes(x = index, y = y)) +
       geom_flower(aes(fill = y), shape=i, colour = NA) +
       coord_polar() +
       theme_void() +
       ggtitle(i) + 
       scale_fill_gradient(low="yellow", high="darkorange", 
                           guide = 'none')
    )() |>
    print()
}
```
